### 2014-04-10 paper/designspace.tex
many paper updates.  need to finish print of whistle and resistors, and take photos, then we are done????

\section{The Design Space of Pipes}
\bjoern{would be nice to have a bridge from related work to here.}
For the purposes of interaction desing, we have identified four relevant dimensions of pipe design: openings, path constraints \bjoern{note the rename - replicate in figure}, topologies, and inserted media.  We describe the space below.

\begin{figure}[t]
\centering
    \includegraphics[width=1.0\columnwidth]{figures/tubespace.pdf}
\caption{The design space of pipes.  See text for details.}
\label{fig:pipespace}
\end{figure}

\subsection{Openings}
Pipes inside solid objects can be either open to the outside on both ends, semi-enclosed, or fully enclosed (see Figure \ref{fig:pipespace}). For interactive devices, openings can be either user-facing (in places a user would for example touch a model); or system-facing, where sensors and actuators may be connected.
%\george{difference between open/return not clear}
%Pipes can have four opening styles: open, return, semi-closed, and fully enclosed .  As we are interested in interactive devices, we distinguish a ``user side"---part of the surface or interior of an object facing its user for interaction; and a ``system side"---where other hardware such as electronics may be connected.  We refer to a pipe which is ``open on the user side'' to have its contents \emph{physically accessible} to a user \emph{at interaction time}.  ``Closed on the user side'' implies physically cut off from the user.  Similarly with the system side.  User or system manipulation of pipe contents are possible even without physical access: e.g., via soundwaves. \george{confusing}

\begin{figure}[t]
\centering
    \includegraphics[width=1.0\columnwidth]{figures/types.png}
\caption{Pipe openings: A shows a pair of open tubes filled with copper paint to power an LED.  B shows a return pipe with EL wire threaded through it.  C shows two semi-closed pipes capped by rubber membranes.  In D, a rabbit has a fully enclosed chamber to modify its weight.}
\label{fig:openings}
\end{figure}

Open pipes may be used to create electronic circuitry.  For example, an open pipe filled with conductive paint and connected to a battery can provide power to an LED (see Figure \ref{fig:openings}A).
%
Semi-enclosed pipes are capped on one side. Caps can for example be used to create tactile output. By fabricating a cap from a malleable material and attaching a pump to the open ended, haptic feedback can be generated by raising and depressing the flexible cap (see Figure \ref{fig:openings}C).
%
A fully enclosed pipe or cavities are entirely internal to the object.  Fully enclosed pipes may be used as weight modifiers~\cite{Prevost-makeitstand} and for object identification~\cite{Willis-infrastructs}, as in Figure \ref{fig:openings}D.  They can also contain for water or particles that would otherwise fall out.\bjoern{how would you get the particles inside?}

\subsection{Path Constraints}
Pipe designs may require specific start and end points, but leave the precise path flexible. For example, circuit schematics specify connectivity, but many routing solutions are valid. Alternatively, the shape of the path itself may be important. For example, in neon signs, single tubes are bent into letterforms. Threading
 electroluminescent (EL) wire through a clear pipe enables makers to emulate neon art (see Figure \ref{fig:openings}B). 
In a maze (see Figure \ref{fig:maze}), both specific terminals as well as a specific path (the maze itself) matter. 
% Alternatively, the neon sign's output in Figure \ref{fig:neon} is based only on the \emph{path} of the pipe.  Focused on just the exterior features is the touch sensitive toy in Figure \ref{fig:toys}, where pipes must exit the toy at the frontal lobe and parietal lobes to make those \emph{specific locations} touch-sensitive.  A pipe for weight modification does require specific terminals or a specific path (see Figure \ref{fig:openings}D).

\subsection{Pipe Topologies}
Pipe network topologies enable splitting or mixing of media (see Figure \ref{fig:pipespace}).  Star and tree topologies extend the splitting and mixing primitives, encompassing multiple-in and multiple-out configurations; we use star topologies to create different conductive paths in touch-sensitive toys, with a single terminal for swept-frequency capacitive sensing (Figure~\ref{fig:toys}).

\subsection{Media in Pipes}
After printing, different media can be inserted into pipes to change affordances and capabilities. We consider gas, liquids, solids, particulates, and threadables. 
%
Compressible fluids (``gas'') and incompressible fluids (``liquids'')  inside semi-enclosed pipes can create haptic feedback; pressure can also be sensed to act as an input~\cite{Slyper-shape}. Gases can be used as carriers for scents or fog. Conductive liquids like copper paint, when dried, turn pipes into conductive paths and can also fix electronic components in place.
%
%got to here
Filling pipes with solids at print time enables designers to exploit the difference in material properties between pipe and contents. For example, in Printed Optics~\cite{Willis-printedoptics}, different refractive indices between solid cladding material and transparent material inside pipes made light transport along pipes for input and output possible.
%
Particulates, either printed in-place or inserted, can be of varying densities and can contain particles of various and variable size.  A single large particle can be used for display.  Small particles in fluid can provide variable stiffness haptics via jamming \cite{Follmer-jamming}. \bjoern{I thought jamming is about particles in a vacuum? why liquid?}
%
Threadable media, such as regular wire, electroluminescent wire, or commercial fiberoptic cables, can be threaded through pipes post-printing. Threading such media enables addition of conduction, illumination, or optical sensing using materials that can easily be procured, but cannot themselves be printed, e.g., on consumer printers with a single material. 
% This empowers inexpensive printers: for example, a Printed Optics-style interface can be created on a consumer-grade 3D printer with pipes and fiber optic cable.

We next show how different pipe openings, constraints and topologies can be created with the help of a new design tool, \systemname.

>>>>

\section{The Design Space of Pipes}
Our review of prior art indicates the potential benifits of using interior pipes for adding interactivity to 3D printed objects, although no general purpose tools currently exist for this purpose. To guide the design of such tool, we now provide a deisgn space of the types of pipes that a maker may want to utilize. We have identified four relevant dimensions of pipe design: \emph{openings}, \emph{path constraints}, \emph{topologies}, and inserted \emph{media}.  We describe the space below.

\begin{figure}[t]
\centering
    \includegraphics[width=1.0\columnwidth]{figures/tubespace.pdf}
\caption{The design space of pipes.  \emph{Pipe openings}, \emph{path constraints}, \emph{topologies}, and inserted \emph{media} give rise to unique input/output possibilities.}
\label{fig:pipespace}
\end{figure}

\subsection{Openings}
Pipes inside solid objects can be either \emph{open} to the outside on both ends, \emph{semi-enclosed}, or \emph{fully enclosed} (see Figure \ref{fig:pipespace}). For interactive devices, openings can be either user-facing (in places a user would for example touch a model); or system-facing, where sensors and actuators may be connected.
%\george{difference between open/return not clear}
%Pipes can have four opening styles: open, return, semi-closed, and fully enclosed .  As we are interested in interactive devices, we distinguish a ``user side"---part of the surface or interior of an object facing its user for interaction; and a ``system side"---where other hardware such as electronics may be connected.  We refer to a pipe which is ``open on the user side'' to have its contents \emph{physically accessible} to a user \emph{at interaction time}.  ``Closed on the user side'' implies physically cut off from the user.  Similarly with the system side.  User or system manipulation of pipe contents are possible even without physical access: e.g., via soundwaves. \george{confusing}

\begin{figure}[t]
\centering
    \includegraphics[width=1.0\columnwidth]{figures/types.png}
\caption{Pipe openings: A shows a pair of \emph{open} pipes filled with copper paint to power an LED.  B shows an \emph{open} pipe with EL wire threaded through it.  C shows two \emph{semi-closed} pipes capped by rubber membranes.  In D, a rabbit has a \emph{fully enclosed} chamber to modify its weight.}
\label{fig:openings}
\end{figure}

\emph{Open} pipes may be used to create electronic circuitry.  For example, an open pipe filled with conductive paint and connected to a battery can provide power to an LED (see Figure \ref{fig:openings}A).
%
\emph{Semi-enclosed} pipes are capped on one side and could be used to create create tactile output. By fabricating a cap from a malleable material and attaching a pump to the open ended, haptic feedback can be generated by raising and depressing the flexible cap (see Figure \ref{fig:openings}C).
%
\emph{Fully enclosed} pipes (cavities) are entirely internal to the object.  Fully enclosed pipes may be used as weight modifiers~\cite{Prevost-makeitstand} and for object identification~\cite{Willis-infrastructs}, as in Figure \ref{fig:openings}D.  If printed in parts and assembled post-print, enclosed pipes could also contain water or particles that would otherwise fall out.

\subsection{Path Constraints}
Depending on the purpose of the pipe, its geometry may be constrained to specific locations within the 3D model. An \emph{endpoint-constrained} pipe requires specific start and end points, but the specific route of the internal path is irrelevant. This would be the case for circuit schematics, where the electronic components have a fixed location, but many routing solutions are valid. In contrast, with a \emph{path-constrained} pipe the shape of the path is important. For example, threading electroluminescent (EL) wire through a clear pipe enables makers to emulate neon art (see Figure \ref{fig:openings}B). A \emph{fully constrained} pipe requires both specific terminals as well as a specific path. If a maker wishes the EL wire to exit the rear of the sign to hide the electronics, it would require a fully-constrained path (Figure \ref{fig:neon}). Finally, an \emph{unconstrained} pipe does not need a specific path or terminal points. Using a cavity to modify an object's weight is an example of an unconstrained pipe.

\subsection{Pipe Topologies}
Pipe network topologies enable \emph{splitting} or \emph{mixing} of media (see Figure \ref{fig:pipespace}).  \emph{Star} and \emph{tree} topologies extend the splitting and mixing primitives, encompassing multiple-in and multiple-out configurations; we use star topologies to create different conductive paths in touch-sensitive toys, with a single terminal for swept-frequency capacitive sensing (Figure~\ref{fig:examples}a).

\subsection{Media in Pipes}
After printing, different media can be inserted into pipes to change affordances and capabilities. We consider \emph{gas}, \emph{liquids}, \emph{solids}, \emph{particulates}, and \emph{threadables}. 
%
Compressible fluids (\emph{``gas''}) and incompressible fluids (\emph{``liquids''})  inside semi-enclosed pipes can create haptic feedback; pressure can also be sensed to act as an input~\cite{Slyper-shape}. Gases can be used as carriers for scents or fog. Conductive liquids like copper paint, when dried, turn pipes into conductive paths and can also fix electronic components in place.
%
Filling pipes with \emph{solids} at print time enables designers to exploit the difference in material properties between pipe and contents. For example, in Printed Optics~\cite{Willis-printedoptics}, different refractive indices between solid cladding material and transparent material inside pipes allowed light transport along pipes for input and output.
%
\emph{Particulates}, either printed in-place or inserted, can be of varying densities and can contain particles of various and variable size.  A single large particle can be used for display.  Small particles in variable amounts of fluid can provide variable stiffness haptics via jamming \cite{Follmer-jamming}.
%
\emph{Threadable} media, such as regular wire, electroluminescent wire, or commercial fiberoptic cables, can be threaded through pipes post-printing. Threading such media enables addition of conduction, illumination, or optical sensing using materials that can easily be procured, but cannot themselves be printed, e.g., on consumer printers with a single material. 
% This empowers inexpensive printers: for example, a Printed Optics-style interface can be created on a consumer-grade 3D printer with pipes and fiber optic cable.

In the next section we how different pipes within this design space space can be created with the help of our new design tool, \systemnamenospace.

### 2014-04-10 paper/discussion.tex
many paper updates.  need to finish print of whistle and resistors, and take photos, then we are done????

\begin{figure}[h]
\centering
    \includegraphics[width=3.4in]{figures/placeholder/direct.jpg}
\caption{By attaching a servo to a gate structure, it is possible to direct fluid through that pipe.  The gate not only has binary positions (a) and (b), but can direct fractional flow, as well (c).  These could be used, for example, in science exhibits.}
\label{fig:direct}
\end{figure}

@@

\begin{figure}[h]
\centering
    \includegraphics[width=3.4in]{figures/placeholder/helmholz.jpg}
\caption{This box (a) contains an enclosed Helmholz resonator, seen in the cutaway view (b).  This resonator creates a sound when air is blown through the pipe at the top, because air compresses at the small neck joint between the chamber and the pipe.  \valkyrie{need to get a test print of this that actually works}}
\label{fig:ocarina}
\end{figure}

\emph{Haptic} outputs are possible through compressible and incompressible fluids.  Fluids in pipes capped with flexible membranes can actuate particulates and create programmable pliability as in Jamming User Interfaces \cite{Follmer-jamming}. \bjoern{Didn't we describe the first part already? Also I'm still skeptical about jamming, maybe skip it.}

@@

\subsection{Identification}
\bjoern{this section should be about half as long as it is right now.}
In addition to active inputs from users, pipes can be used for passive \emph{identification}.  Fully-enclosed pipes which ar enot visually apparent can change that weight, acoustic resonance, and capacitive signature of a 3D object.

Each of these properties could be used for object identification. As seen in Touch \& Activate \cite{Ono-touchandactivate}, different objects have different acoustic signatures.  As such, two objects that are visually identical but have differing internal chamber geometry on their interior may be distinguished acoustically.  Thus we can recall an object's identity once its acoustic signature has been recorded.  Similarly, the Touch\'{e} system \cite{Sato-touche} relies on distinguishing capacitive signatures of objects, both in static states and as their boundary conditions are changed by human interaction. Weight could also be used as a simple metric of object identity, but would be less distinctive than acoustic or capacitive signatures.

@@

\begin{figure}[h]
\centering
    \includegraphics[width=3.4in]{figures/electronics.png}
\caption{(a) A capacitor we manufactured on a 3D printer and coated, post-print, with copper paint.  This particular one holds a charge of xx.  (b) Comparison of theoretical printed resistors.  For a cylindrical wire, resistance = $\rho\frac{l}{\pi r^2}$, where $\rho$ is based on the conductivity of the medium.  In our case, assuming $ r << l$, resistance(B) < resistance(C) < resistance(A).  (c) A possible design for a cell phone antenna augmentation inside a cell phone case. \bjoern{skip the antenna since we haven't investigated this. If you have the capacitor and you're confident you know how to measure its charge, add it. The different conductors with different measurements are easy to do and worthwhile. Maybe combine these two with the helmholtz resonator into one smaller figure.}}
\label{fig:electronics}
\end{figure}

@@

\valkyrie{do we want those figures of the helmholz thing and the servo redirection gate?}

>>>>

\begin{figure}[h]
\centering
    \includegraphics[width=3.4in]{figures/gates.png}
\caption{By attaching a servo to a gate structure, it is possible to direct fluid through that pipe.  The gate not only has binary positions (a) and (b), but can direct fractional flow, as well (c).  This could be used, for example, in science exhibits.}
\label{fig:direct}
\end{figure}

@@

\begin{figure}[h]
\centering
    \includegraphics[width=3.4in]{figures/placeholder/helmholz.jpg}
\caption{Some possible pipe uses.  The printed whistle in (a) is a Helmholz resonator: it creates a sound when air is blown through the pipe at the front, because air compresses at the small neck joint between the internal chamber and the pipe.  In (b), a capacitor manufactured on a 3D printer coated post-print with copper paint.  This particular one holds a charge of \valkyrie{XX}.  (c) gives a comparison of pipe-based resistors, also filled with copper paint.  For a cylindrical wire, resistance = $\rho\frac{l}{\pi r^2}$, where $\rho$ is based on the conductivity of the medium.  In our case, because $r << l$, $resistance(B) < resistance(C) < resistance(A)$.}
\label{fig:ocarina}
\end{figure}

\emph{Haptic} outputs are possible through compressible and incompressible fluids, as discussed, or through threadable inputs: e.g., muscle wire can be inserted in flexible pipes and actuated, similar to shape changing interfaces \cite{Coelho-shapechanging}.

@@

\subsection{Identification}
In addition to active inputs from users, pipes can be used for passive \emph{identification}.  Fully-enclosed pipes which are not visually apparent can change the weight, acoustic resonance, or capacitive signature of a 3D object.

Each of these properties could be used for object identification. As seen in Touch \& Activate \cite{Ono-touchandactivate}, different objects have different acoustic signatures.  As such, two objects that are visually identical but have differing internal chamber geometry on their interior may be distinguished acoustically.  Thus we can recall an object's identity once its acoustic signature has been recorded.  Similarly, the Touch\'{e} system \cite{Sato-touche} relies on distinguishing \emph{capacitive} signatures of objects, both in static states and as their boundary conditions are changed by human interaction. Weight could also be used as a simple metric of object identity, but would be less distinctive than acoustic or capacitive signatures.

@@

@@

### 2014-04-10 paper/exampleobjects.tex
many paper updates.  need to finish print of whistle and resistors, and take photos, then we are done????

\section{Example Objects}
\tovi{it might be intersting to relate our examples to existing HCI papers - in order we have: Touche, Dynamically Changeable Physical Buttons/PneUI, Midas, Suaron. I guess the maze doesnt really relate to anything.}
\tovi{UIST is someitmes cranky about providing enough details so other can reporduce the work. We might want to be more careful about providing specific part numbers of specifications.}
\bjoern{subsection headings need to be updated in light of the simplification and label changes in the design space.}
To evaluate and highlight our tool's capabilities, we fabricated a set of five prototype objects designed using \systemnamenospace.  \tovi{might want to specify the exact printer models we used here, and say we used multipel printers to show that are approach is printer independant.}
%All prototypes were fabricated in a single piece, unless described otherwise. \bjoern{2 of 5 will have been fabricated otherwise -  Radio, Neon sign - both in pieces.}

\subsection{Touch-sensitive Toys (open, liquid, star, terminals)}

To show how conductive fluids can be used for touch sensing, we created a set of touch-sensitive toys and a companion app that identifies the objects as well as touched parts of the objects (inspired by \cite{Harrison-acoustic}). The objects --- a brain and boat, in our example --- can be set on a base, upon which a speaker announces ``brain'' or ``boat''. Touching the brain model's front protrusion yields the announcement ``olfactory bulb''. The distinct touch points on each object are connected by an interior star topology, and touch sensing is performed via a single conductive connection to the base, using swept-frequency capacitive sensing~\cite{Sato-touche} (see Figure \ref{fig:toys}). %We built a smart base which can distinguish between the toys and also determine which toy is mounted:
Since each toy and each gesture has a distinct capacitive signature, we use a simple classifier trained to detect both toy and gesture based on profile.

The models we used for the toys were downloaded from the  Thingiverse website. We then used \systemnamenospace to create the desired internal pipes. All components were fabricated on a Makerbot, support-free.  After printing, we injected copper paint (CuPro-Cote) into the interior pipes using a craft syringe.  This paint requires approximately 2 days to fully dry in this configuration (3mm diameter tubes, maximum tube length 10cm).  Our smart base is powered by an Arduino Uno running open-source SFCS code\footnote{http://www.instructables.com/id/Touche-for-Arduino-Advanced-touch-sensing}. An advatnage of using this paint is that it also serves as an adhesive, so no soldering was required. \tovi{confirm}

\subsection{Breathing Bunny (semi-closed, gas, terminals)}

To demonstrate how gases can be used to both provide sensation to the user through openings and to deform a model internally, we created a rabbit with a pair of tubes that can simulate breathing (see Figure \ref{fig:breathe}).  When the rabbit inhales through its nose, its abdomen rises, and as it exhales its abdomen falls.  For this, we used a combination air/vacuum pump: one terminal creates a vacuum while the other creates positive pressure.  
The rabbit model was downloaded from the Thingiverse website. \tovi{confirm} We then used \systemnamenospace to add two pipes: one open pipe exiting at its mouth, and one semi-closed pipe capped at its abdomen.  We connected one pipe to each of our pump's terminals, and using a programmable power supply we mimic a rabbit's breathing pattern.  This example was printed on the Objet using a flexible material and its support material was flushed post-print.\tovi{give pump part number/details}

\subsection{Custom Radio (open, threadable, terminals)}
\tovi{stick with the strucutre - describe what it does, then describe modeling and fabrication (how it was built)}
A custom radio shows how pipes can be used to integrate electronic components into the user-facing surface of an object. The radio includes a power and volume dial, and LED which goes on when the radio is on, and a tuner dial. \tovi{confirm}. A Arduino Pro Micro microcontroller, Si4703 radio tuner, and  [x volt] power supply are embedded in the base of the radio. The sound is played through a single speaker.

The radio was manually modelled in SolidWorks. We included slots on the 3D model matching the shape of the electronic components that were to be embedded. The model was then imported into \systemnamenospace, where we created a network of 8 individual open pipes (3 per potentiometer, 1 for the LED, 1 for the speaker) to connect the the comments to the base of the radio. We fabricated the radio itself on our Makerbot, support-free, but cut in two pieces to allow the Arduino and battery pack to be inserted. We then threaded the pipes with wires to connect the compoenents to the microcontroller at the base of the radio.

%This design with 8 pipes (3 per potentiometer, 1 for the LED, 1 for the speaker) showed the limitations of our incremental routing strategy: since we voxelize and remesh for each pipe, surface detail is incrementally lost. The problem can be addressed by increasing the voxel resolution. At 512x512x512 voxels, surface geometry is not notably degraded. However, a single cut operation takes around 3 minutes \bjoern{on what hardware? mention earlier}, compared to near-instant performance on a $128^3$ voxel grid.  We fabricated the radio itself on our Makerbot, support-free but cut in two pieces to allow the Arduino and battery pack to be inserted. TOVI: Move to discussion if you want to keep this text.

\bjoern{I ran out of steam somewhere here...}
\tovi{relate this to sauron as well}
\subsection{Presence-aware Pen Holder (open, solid, ternimals)}
\bjoern{isn't this a threadable instead of a solid?} \bjoern{the model looks like the acutal pipes are really short and basically straight at the bottom of the model. This may raise the question if one couldn't just use the line sensors without any fiber optic cable in the middle.}
Our presence-aware pen holder can distinguish which tool or tools a user has picked up (see Figure \ref{fig:pens}).  Our pen holder uses a modification of the FlyEye technique described by Wimmer in \cite{Wimmer-flyeye} in which \bjoern{describe...}.

The model was created/downloaded from [x]. We used \systemnamenospace to add four internal open pipes connecting each pen chamber to the based on the pen holder. This prototype was built by our Objet and support was flushed post-print.\bjoern{Incorrect. Looks like a Makerbot print.} \valkyrie{this whole paragraph is unclear}

Each pipe was then threaded with a fiber optic cable.  We use a single cable per pipe; our 6mm diameter fiber optic cables, in comparison to the fine tubes used in the original work , can send and receive through the same cable. \bjoern{is this because it has many strands?}  At the base of each tube is a QRE1113 line sensor digital breakout board, which has an integrated IR emitter and receiver.   When a pen is in its appointed place, the emitted infrared light is reflected off its bottom and travels back to the receiver, where it registers as bright.

\subsection{Animated Neon Sign (return, threadable, path)}

\begin{figure}[h!]
\centering
    \includegraphics[width=3.4in]{figures/sign.png}
\caption{A two-state animated neon sign designed using our tool.  (a) shows the input SVG files, with the two states of the mouth drawn in different colors. \bjoern{it's hard to tell these colors apart!}  (b) shows the resultant mesh when we subtracted the generated tubes. \bjoern{hard to see what's going on here, too.} In (c), we show the fabricated sign.}
\label{fig:neon}
\end{figure}

Neon art is traditionally made from hand-formed glass tubes containing neon gas.  The tubes light up when a current is passed through them.  For this type of art, the path of the tubes is of crucial importance, as it determines how the sign will look.  We designed a custom neon robot head which can be animated to ``talk''.  
\tovi{in the system seciton do we ever discuss how entry and exit points are specified for path-constrained pipes?}
The geomertry of the model was imported into \systemnamenospace, which created the assocaited Pipes using its internal path routing tool. We then printed the model on the Objet using a transparent material. The model was printed in two halves, cutting through the plane of pipe path, to facilitate assembly.

The pipes were threaded with electroluminescent wire which is lit in sequence to create an animation using an EL wire sequencer. Pipes enter and exit through the rear to hide the EL wire controls.

\subsection{Maze (fully enclosed, particulate, tree, terminals + path)}

Our final example object is a maze game with a marble that can be navigated through a maze-like path, with exits on either side of the model.

The maze is based on an SVG file we manually created and processed with our internal path routing tool.  The Maze was printed on the Objet using a transparent material. We created the maze in two halves that were fastened together via glue, to support the removal of support materials using a high-pressure water tool.  \bjoern{check - i think the latest maze is in fact done with our tool? Will need rewriting.}

>>>>

\section{Example Objects}
To evaluate and highlight our tool's capabilities, we fabricated a set of five prototype objects designed using \systemnamenospace.  To show that our approach is printer-independent, we fabricated some on a Makerbot Replicator 2 and the rest on an Objet Connex 260.  Each prototype object explores the expansion of an existing sensing or actuation technique to the surface of 3D printed objects.
%All prototypes were fabricated in a single piece, unless described otherwise. \bjoern{2 of 5 will have been fabricated otherwise -  Radio, Neon sign - both in pieces.}

\subsection{Touch-sensitive Toys (open, liquid, star, endpoint)}

To expand Touch\'{e} \cite{Sato-touche} to 3D printed forms, we created a set of touch-sensitive toys and a companion app that identifies the objects as well as touched parts of the objects (inspired by \cite{Harrison-acoustic}). The objects --- a brain and boat, in our example --- can be set on a base, upon which a speaker announces ``brain'' or ``boat''. Touching the brain model's front protrusion yields the announcement ``olfactory bulb''. The distinct touch points on each object are connected by an interior star topology, and touch sensing is performed via a single conductive connection to the base, using swept-frequency capacitive sensing (see Figure \ref{fig:examples}a). %We built a smart base which can distinguish between the toys and also determine which toy is mounted:
Since each toy and each gesture has a distinct capacitive signature, we use a simple classifier trained to detect both toy and gesture based on profile.

The models we used for the toys were downloaded from the  Thingiverse website. We then used \systemname to create the desired internal pipes. All components were fabricated on a Makerbot, support-free.  After printing, we injected copper paint (CuPro-Cote) into the interior pipes using a craft syringe.  This paint requires approximately 2 days to fully dry in this configuration (3mm diameter tubes, maximum tube length 10cm).  Our smart base is powered by an Arduino Uno running open-source SFCS code\footnote{\url{http://www.instructables.com/id/Touche-for-Arduino-Advanced-touch-sensing}}. \valkyrie{removed sentence about solderless paint: all active components are affixed to a breadboard, and the only thing actually attached to the paint is a wire (which we plug into the breadboard).  granted, we did not have to solder the wire, but it seems weird to mention that...?} %An advantage of using this paint is that it also serves as an adhesive, so no soldering was required. \tovi{confirm}

\subsection{Breathing Bunny (semi-closed, gas, endpoint)}

With inspiration from PneUIs \cite{Yao-pneui} for using gases to both provide sensation to the user through openings and deform a model internally, we created a rabbit with a pair of tubes that can simulate breathing (see Figure \ref{fig:examples}b).  When the rabbit inhales through its nose, its abdomen rises, and as it exhales its abdomen falls.  For this, we used a combination air/vacuum pump (ROB-10398): one terminal creates a vacuum while the other creates positive pressure.  
The rabbit model, the well-known Stanford Bunny\footnote{\url{http://en.wikipedia.org/wiki/Stanford_bunny}}, was downloaded from the Thingiverse website. We then used \systemname to add two pipes: one open pipe exiting at its mouth, and one semi-closed pipe capped at its abdomen.  We connected one pipe to each of our pump's terminals, and using a programmable power supply we mimic a rabbit's breathing pattern.  This example was printed on the Objet using a flexible material and its support material was flushed post-print.

\subsection{Custom Radio (open, threadable, endpoint)}
A custom radio (figure \ref{fig:examples}c) shows how pipes can be used to integrate electronic components into the user-facing surface of an object. The radio includes a power and volume dial, and LED which goes on when the radio is on, and a tuner dial. An Arduino Pro Micro microcontroller, Si4703 FM tuner, and 6V (4xAA) power supply are embedded in the base of the radio. The sound is played through a single .5 Watt speaker. This shows a Midas-like \cite{Savage-midas} possibility: multiple terminals all routed back to a shared endpoint (in this case, the Arduino).

The radio shell was modelled in SolidWorks. We included slots on the 3D model matching the shape of the electronic components that were to be embedded. The model was then imported into \systemnamenospace, where we created a network of 8 individual open pipes (3 per potentiometer, 1 for the LED, 1 for the speaker) to connect the the comments to the base of the radio. We fabricated the radio itself on our Makerbot, support-free, but cut in two pieces to allow the Arduino and battery pack to be inserted. We then threaded the pipes with wires to connect the compoenents to the microcontroller at the base of the radio.

\subsection{Presence-aware Pen Holder (open, threadable, endpoint)}
\bjoern{the model looks like the acutal pipes are really short and basically straight at the bottom of the model. This may raise the question if one couldn't just use the line sensors without any fiber optic cable in the middle.}  \valkyrie{I suspect either would work; we don't have to use this as a threadable example if we don't want to, since the radio and the EL wire are both threadable already}
Our presence-aware pen holder can distinguish which tool or tools a user has picked up (see Figure \ref{fig:examples}d).  Our pen holder uses a modification of the FlyEye technique described by Wimmer \cite{Wimmer-flyeye} in which an IR camera senses a user's grasp on an object, getting data from IR light transmitted on fiber optic cables.  We marry this technique here with the Sauron technique \cite{Savage-sauron}: instead of using mirrors to redirect light and motion sensing as Sauron does, we use fiber optic cables.

The penholder body was created in Solidworks. We used \systemname to add four internal open pipes connecting each pen chamber to the based on the pen holder. This prototype was built by our Makerbot support-free.  Each pipe was then threaded with a fiber optic cable.  We use a single cable per pipe; our 6mm diameter fiber optic cables, in comparison to the fine tubes used in the original work, can send and receive through the same cable because their cross-sectional area is large enough to accommodate both functions. At the base of each tube is a QRE1113 line sensor digital breakout board, which has an integrated IR emitter and receiver.   When a pen is in its appointed place, the emitted infrared light is reflected off its bottom and travels back to the receiver, where it registers as bright.

\subsection{Maze (fully enclosed, particulate, tree, path)}

We created a maze game with a marble that can be navigated through a maze-like path, with exits on either side of the model (Figure \ref{fig:examples}e).

The maze is based on an SVG file we manually created and processed with our internal path routing tool.  We printed the maze on the Objet using a transparent material. The final artifact was fabricated in two halves fastened together via glue, to allow removal of support material with a high-pressure water tool.  We plugged the ends of the tubes post-print with additional material to ensure the ball does not escape.

\subsection{Animated Neon Sign (open, threadable, fully constrained)}

\begin{figure}[t]
\centering
    \includegraphics[width=3.4in]{figures/sign.png}
\caption{A two-state animated neon sign designed using our tool.  (a) shows the input SVG files, with the eyes (non-animated) in black and the two states of the mouth in {\color{blue}blue} and {\color{gray}gray}.  (b) shows the resultant mesh when we subtracted the generated tubes, with rear exit tubes circled in {\color{blue}blue}. In (c), we show the fabricated sign.}
\label{fig:neon}
\end{figure}

Neon art is traditionally made from hand-formed glass tubes containing neon gas.  The tubes light up when a current is passed through them.  For this type of art, the path of the tubes is of crucial importance, as it determines how the sign will look.  We designed a custom neon robot head which can be animated to ``talk''. 
The external geometry of the model was imported into \systemnamenospace, which created the associated pipes using its internal path routing tool.  We added additional endpoint-constrained tubes as exits out the rear of the model (see Figure \ref{fig:neon}B).  We then printed the model on the Objet using transparent material. The model was printed in two halves, cutting through the plane of internal path, to facilitate assembly.

The pipes were threaded with electroluminescent wire which is lit in sequence, using an EL wire sequencer, to create the animation. Pipes enter and exit through the rear to hide the EL wire controls.

### 2014-04-10 paper/introduction.tex
many paper updates.  need to finish print of whistle and resistors, and take photos, then we are done????

We describe a new design space of pipes and hollow chambers for the purpose of interaction design, where variables include openings, topologies, and inserted media. By exploring this design space new opportunities for adding interactivity to 3D printed objects can be realized. For example, injecting conductive paint can turn pipes into conductors, which can in turn be used to integrate electronic components into prints. Electroluminescent (EL) wire can be threaded through pipes to create computer-controlled visual output. And air can be pumped through pipes to produce haptic effects.  Existing sensing and actuation approaches, such as FlyEye \cite{Wimmer-flyeye}, Swept Frequency Capacitive Sensing \cite{Sato-touche}, and Jamming User Interfaces \cite{Follmer-jamming}, can be enhanced by the use of pipes: makers utilizing such I/O strategies can locate input and output at arbitrary points on a printed 3D object's surface.  \tovi{Is enhanced the right word? Or would "enabled" be better - since these techniques can now be applied to 3D printed objects} \valkyrie{I like enhanced; we are allowing them to be applied in new places}  We leverage these existing techniques for sensing and actuation while our work's novelty is in \emph{internal modeling} for 3D printing and the \emph{redirection} of I/O to arbitrary locations on a device.

>>>>

We describe a new design space of pipes and hollow chambers for the purpose of interaction design, where variables include openings, topologies, and inserted media. By exploring this design space new opportunities for adding interactivity to 3D printed objects can be realized. For example, injecting conductive paint can turn pipes into conductors, which can in turn be used to integrate electronic components into prints. Electroluminescent (EL) wire can be threaded through pipes to create computer-controlled visual output. And air can be pumped through pipes to produce haptic effects.  Existing sensing and actuation approaches, such as FlyEye \cite{Wimmer-flyeye}, Swept Frequency Capacitive Sensing \cite{Sato-touche}, and Jamming User Interfaces \cite{Follmer-jamming}, can be expanded to new contexts by the use of pipes: makers utilizing such I/O strategies can locate input and output at arbitrary points on a printed 3D object's surface.  We leverage these existing techniques for sensing and actuation while our work's novelty is in \emph{internal modeling} for 3D printing and the \emph{redirection} of I/O to arbitrary locations on a device.

### 2014-04-10 paper/limitations.tex
many paper updates.  need to finish print of whistle and resistors, and take photos, then we are done????

>>>>

This basic approach of using a discrete rod for physical simulation does have various limitations, in particular it
cannot recover if the rod becomes tangled or unstable. 
However we have observed these problems only when attempting to break
the system; they did not occur in any of the cases shown in our figures.
Another issue is parameters, as the penalty forces used in PBD do not
correspond to physical units. We avoided serious issues here by uniformly
scaling the initial mesh to a unit box, and tuning the parameters on several test cases.
